Universal Anonymous Signatures

### 2022-08-05 prism-backend/research/gsac/app-uas-proofs.tex
Updating proofs.

\begin{figure}[ht!]
  \centering  
  \scalebox{0.9}{
    \procedure{$\Exp^{\UAS-acc-b}_\adv(1^\secpar)$}{%
      \parm \gets \Setup(1^\secpar) \\
      \pcfor d\in\bin: \\
      \pcind (\upk_d,\usk_d) \gets \KeyGen(\parm) \\
      \pcind (\st,\attrs, \lbrace (\cidi_d,\cidu_d,\cattrs_{d,i}) \rbrace_{i\in[n]})
      \gets \adv(\st,\upk_0,\upk_1) \\
      \yissue^* \gets \fissue(\upk_b,
      \lbrace (\cidi_b,\cattrs_{b,i}) \rbrace_{i\in[n]},
      \attrs) \\
      b^* \gets \adv(\st,\yissue*) \\
      \pcreturn b=b^*
    }
  }
  \label{fig:uas-acceptable}
  \caption{Formal definition of \UAS-acceptable \fissue functions.}
\end{figure}

Next, we define the \SimSetup, \ExtractIssue, \ExtractSign functions. \SimSetup
(formally defined in \figref{fig:helper-funcs}).essentially replaces the
\Setup algorithm of each NIZK with the corresponding \NIZKSimSetup algorithm,
and returns the generated parameters along with the
trapdoors. \ExtractIssue and \ExtractSign simply run the \Extract functions
of the corresponding \NIZK scheme, leveraging the trapdoor produced during
\SimSetup. Note that, due to the zero-knowledge and simulation extractability
properties (respectively) of the underlying NIZKs, the result of \SimSetup
is indistinguishable from that of \Setup, and the extracted values match those
used as witnesses in the proofs.

@@

\begin{proof}[\thmref{thm:issue-anonymity-uas}; Issuance anonymity of \CUASGen]
  % We prove the special case in which the adversary makes only one query to the
  % \OBTCHALb oracle. The generalization to polynomially many queries follows from
  % a standard hybrid argument.

  First, observe that, excluding the \yissue value output revealed to the issuer
  during the interactive issuance protocol, \CUASGen essentially runs an
  interactive signing protocol of an \SBCM scheme, augmented with a
  zero-knowledge NIZK for $\NIZKRel_{\Issue}$. As stated in \appref{sapp:sbcm},
  blindness of the (non-augmented) \SBCM scheme, and the zero-knowledge property
  of the NIZK for $\NIZKRel_{\Issue}$, ensure that the challenge credentials are
  indistinguishable between thesmselves. Concretely, we can replace the NIZK
  proofs sent by the user by simulated proofs.

  Finally, note that, to avoid trivial wins by the adversary, the issuance
  anonymity game requires that \fissue either rejects (returns $\yissue=0$) for
  both challenge users, or it accepts for both (returns $\yissue \neq 0$).
  Moreover, in the latter case, since we require that \fissue functions are
  \UAS-acceptable functions, the resulting \yissue value is hiding with respect
  to the input arguments. Thus, adding \yissue does not allow the adversary to
  distinguish credential issuance protocols.  
\end{proof}

@@

  We start from $G_0=\ExpSigAnonb$.  %
  From $G_0$, we consider $G^0_0 = \ExpSigAnonz$. The challenge sent to the
  adversary is $(\csig_0,\yeval) \gets \Sign(\PRVUK[\cuid_0],\PUBOK[\oid],
  \CRED[\scid_0],\msg,\feval)$, where $\csig_0 = (\pi_0,\Ec_{\yinsp})$, with
  $\pi_0 = \NIZKProve^{\NIZKRel_{\Sign}}(\NIZKcrs_{\Sign},(\msg,\feval,\yeval,
  \ceval,\cinsp,\PUBIK[\scid_0],\widetilde{\Eek},\PUBOK[\oid]),(\PRVUK[\cuid_0],
  \CRED[\scid_0],\attrs_{\scid_0},\yeval^1,\yinsp,r,r'))$, $\ceval = \EEnc
  (\widetilde{\Eek},\yeval^1;r)$, and $\cinsp = \EEnc(\PUBOK[\oid],\yinsp;r')$.
  % 
  Further, we build $G_1^0$ from $G_0^0$ by simulating the proof $\pi_0$. That
  is, in $G_1^0$, $\csig_0 = (\pi_0^s,\ceval,\cinsp)$, where $\pi^s_0 =
  \NIZKSim^{\NIZKRel_{\Sign}}(\NIZKcrs_{\Sign},\NIZKtrap,(\msg,\feval,\yeval,
  \ceval,\cinsp,\PUBIK[\scid_0],\PUBOK[\oid]))$. By zero-knowledgeness
  of $\NIZK^{\Sign}$, $G_1^0$ is indistinguishable from $G_0^0$.

  Similarly, we consider $G_0^1$ and $G_1^1$. That is, $G_0^1 = \ExpSigAnono$,
  where the challenge
  sent to the adversary is $(\csig_1,\yeval) \gets \Sign(\PRVUK[\cuid_1],
  \PUBOK[\oid],\CRED[\scid_1],\msg,\feval)$, where $\csig_1 = (\pi_1,\ceval,
  \cinsp)$, with $\pi_1 = \NIZKProve^{\NIZKRel_{\Sign}}(\NIZKcrs_{\Sign},
  (\msg,\feval,\yeval,\ceval,\cinsp,\PUBIK[\scid_1],\widetilde{\Eek},
  \PUBOK[\oid]),(\PRVUK[\cuid_1],\CRED[\scid_1],\attrs_{\scid_1},\yeval^1,
  \yinsp,r,r'))$, $\ceval = \EEnc(\widetilde{\Eek},\yeval^1;r)$, and
  $\cinsp = \EEnc(\PUBOK[\oid],\yinsp;r')$. As before, build $G_1^1$ from
  $G_0^1$, simulating $\pi_1$. That is, in $G_1^1$, $\csig_1 = (\pi_1^s,\ceval,
  \cinsp)$, where $\pi^s_1 = \NIZKSim^{\NIZKRel_{\Sign}}(\NIZKcrs_{\Sign},
  \NIZKtrap,(\msg,\feval,\yeval,\ceval,\cinsp,\PUBIK[\scid_1],\widetilde{\Eek},
  \PUBOK[\oid]))$. Again, by zero-knowledge of $\NIZK^{\Sign}$, $G_1^1$ is
  indistinguishable from $G_0^1$. Note also that $G_1^1$ and $G_1^0$ are
  indistinguishable, due to the IND-CCA property of the encryption scheme
  (so, the \ceval values in $\csig_0$ and $\csig_1$ are indistinguishable),
  in the  challenge oracle used in the anonymity game, we restrict to
  $\PUBIK[\scid_0] = \PUBIK[\scid_1]$, and the respective \cinsp values encrypt
  the same \yinsp value (and, otherwise, \adv~is not allowed to open it).

@@

\begin{proof}[\thmref{thm:issue-forge-uas}; Issuance unforgeability of \CUASGen]
  We show that the probability that \fissue outputs $0$ is negligible, as well
  as the probability that the extracted \usk is not the one that was used to
  request some of the credentials employed to obtain the credential specified by
  the adversary.
  %
  For this purpose, we define two games, $G_0=\ExpForgeIssue$, and $G_1$, which
  is exactly the same, but where, within the \Setup algorithm, we replace
  $\NIZKSetup^{\Issue}$ with $\NIZKSimSetup^{\Issue}$. Due to zero-knowledgeness
  of $\NIZK^{\NIZKRel_{\Issue}}$, both games are indistinguishable.

  Now, observe that the adversary is required to output a credential
  identifier for which associated entries in \trans and \CRED exist; moreover,
  if such a credential was produced by an issuer, we must have access to those
  entries, as issuers are assumed to be honest.
  %
  Then, given that $\NIZKRel_{\Issue}$ is knowledge extractable (which is implied
  by simulation-extractability), in game $G_1$
  we can apply the \NIZKExtract function, which produces a tuple $(\usk,\scred,
  \attrs_{\scred})$ from $\utrans = (\Ccom,\attrs,\sipk,\cred,\NIZKproof)$.
  %
  Since \NIZKproof is accepted by \ExtractIssue, from the soundness of \NIZK, we
  know that all $\cred \in \scred$ are valid signatures over \usk, and their
  respective $\attrs_{\cred}$. Thus, \IdentifyCred returns $1$ for all $(\usk,
  \attrs_{\cred},\cred)$ tuples. That is, all the credentials in \scred given
  to \fissue belong to the same user, who is the owner of \usk.
  %
  Finally, since issuers are honest, we know that $\ATTR[\cid] = \attrs$ and,
  consequently, $\fissue(\usk,\scred,\ATTR[\cid]) = \fissue(\usk,\scred,\attrs)
  = 1$, due to the soundness of \NIZK.
  %
  \qed
\end{proof}

@@

\begin{proof}[\thmref{thm:sign-forge-uas}; Signature unforgeability of \CUASGen]
  As in \thmref{thm:issue-forge-uas}, we define two games, $G_0=\ExpForgeSign$,
  and $G_1$, which is exactly the same but where, within the \Setup algorithm,
  we replace $\NIZKSetup^{\Sign}$ with $\NIZKSimSetup^{\Sign}$. After
  zero-knowledgeness of $\NIZK^{\Sign}$, both games are indistinguishable.
  %
  Next, we show that an adversary winning $G_1$ can be used to break the
  one-more unforgeability property of \SBCM.

  If the verification at line 4 holds, then $(\msg,\feval,\yeval,\ceval,
  \cinsp,\sipk,\opk,\widetilde{\Eek}) \in \NIZKLang^{\Sign}$. Then:
  %

  \paragraph{(a) \Judge accepts $(\yinsp,\iproof)$.} %
  Simulation extractability of $\NIZK^{\Sign}$ thus ensures that $\cinsp =
  \EEnc(\Eek,\yinsp)$, and since the $(\yinsp,\iproof)$ is generated honestly at
  line 5, then \yinsp is the correct decryption of \cinsp, and  correctness of
  $\NIZK^{\Open}$ ensures that \Judge outputs $1$ at line 6.

@@

  \begin{itemize}
  \item $\yeval^0 = \yeval$, where $(\yeval^0,\cdot) = \feval(\usk,\scred,
    \msg)$, and $\yeval$ is as output by \adv~at step 2.    
  \item $\tyeval^1 = \yeval^1$, where $(\cdot,\yeval^1) = \feval(\usk,\scred,
    \msg)$, and $\tyeval^1$ is as extracted by \ExtractSign at line 7.
  \end{itemize}

  Thus, the probability of \adv~winning at line 9 is $0$.

  \paragraph{(c) The output of \finsp matches the output of \Open.} %
  After (a), the \yinsp value output by \Open is the correct decryption of
  \cinsp. After (b), the $(\yeval^0=\yeval,\yeval^1)$ values output by
  \ExtractSign match the evaluation of \feval. Thus, simulation extractability
  of $\NIZK^{\Sign}$ ensures that $\finsp((\yeval^0,\yeval^1),\usk,\scred,\msg)
  = \yinsp$ and, also, that \yinsp matches the $\yinsp'$ value extracted by
  \ExtractSign. Thus, the probability of \adv~winning at line 10 is $0$.

  \paragraph{(d) All {\cred}s are bound to the same \usk.} %
  $\NIZKRel^{\Sign}$ includes a condition that $\forall \cred \in \scred,
  \SBCMVerify(\ipk_{\cred},\cred,\usk,\attrs_{\cred}) = 1$. Thus, simulation
  extractability of $\NIZK^{\Sign}$ and correctness of \SBCM, ensure that all
  credentials involved in the signature contain \usk as their user key (first)
  attribute. Consequently, \IdentifyCred returns $1$ for all the involved
  credentials, and the probability of \adv~winning at line 11 is $0$.

>>>>

\begin{figure}[ht!]
  \centering  
  \scalebox{0.9}{
    \procedure{$\Exp^{\UAS-acc-b}_\adv(1^\secpar)$}{%
      \parm \gets \Setup(1^\secpar) \\
      (\upk,\usk) \gets \KeyGen(\parm) \\
      (\st,\attrs, \lbrace (\cidi,\cidu,\cattrs_i) \rbrace_{i\in[n]})
      \gets \adv(\upk) \\
      \pcif b=0:
      \yissue^* \gets \fissue(\upk,\lbrace (\cidi,\cattrs_i)
      \rbrace_{i\in[n]},\attrs) \\
      \pcelse: \yissue^* \getr \rngfissue \setminus \lbrace 0 \rbrace \\
      b^* \gets \adv(\st,\yissue*) \\
      \pcreturn b=b^*
    }
  }
  \label{fig:uas-acceptable}
  \caption{Formal definition of \UAS-acceptable \fissue functions.}
\end{figure}

Next, we define the \SimSetup, \ExtractIssue, \ExtractSign functions. \SimSetup
(formally defined in \figref{fig:helper-funcs}) essentially replaces the
\Setup algorithm of each NIZK with the corresponding \NIZKSimSetup algorithm,
and returns the generated parameters along with the
trapdoors. \ExtractIssue and \ExtractSign simply run the \Extract functions
of the corresponding \NIZK scheme, leveraging the trapdoor produced during
\SimSetup. Note that, due to the zero-knowledge and simulation extractability
properties (respectively) of the underlying NIZKs, the result of \SimSetup
is indistinguishable from that of \Setup, and the extracted values match those
used as witnesses in the proofs.

@@

\begin{proof}[\thmref{thm:issue-anonymity-uas}; Issuance anonymity of \CUASGen]
  We prove the special case in which the adversary makes only one query to the
  \OBTCHALb oracle. The generalization to polynomially many queries follows
  from a standard hybrid argument.

  We start from the \ExpIssAnonb, as defined in \figref{fig:exp-uas-anonb}.

  Game 1: We simulate the NIZK proof corresponding to $\NIZKRel_{\Issue}$ that
  the \OBTCHALb oracle produces when queried by \adv. The zero-knowledge
  property of the NIZK system ensures that game 1 is indistinguishable from
  the original experiment.

  Game 2: In queries to \OBTCHALb, before simulating the NIZK proof, we replace
  the produced \yissue value with a random element from \rngfissue. Since
  \fissue is a \UAS-acceptable function, game 2 is indistinguishable from game
  1.

  From an adversary \adv~against Game 2, we can build an adversary \advB against
  \SBCM's blindness property. The simulation that
  \advB does is straightforward. \advB runs Game 2 for \adv, randomly picking
  a number $q < n$, where $n$ is the number of users that \adv~creates via
  \UGEN, and assigning to the $q$-th user the $(\SBCMvk,\SBCMsk)$ key pair
  that \advB receives in its blindness game. Then, when \adv~makes a query
  to its \OBTCHALb oracle, if the chosen issuer is the $q$-th user, \advB
  forwards \adv's messages to its own $\langle COM(\SBCMvk,
  \overline{\smsg}_{\bin},\smsg),\cdot\rangle$ oracle. Finally, \advB outputs
  whatever \adv~does. Note that, since in Game 2 we replace the \yissue value
  with a random value, and simulate the $\NIZKRel_{\Issue}$ proof, \adv~is not
  receiving information about the blindly signed messages beyond what \advB
  forwards from its own oracle. Thus, if \adv~wins in its game, \advB wins in
  its blindness game. However, since \advB has to guess the correct user
  that \adv~will use as issuer in its \OBTCHALb oracle, we have a $1/n$ security
  loss, where $n$ is the total number of users created by \adv.
  %
  \qed
\end{proof}

@@

  We start from $G_0=\ExpSigAnonb$.  %
  From $G_0$, we consider $G^0_0 = \ExpSigAnonz$. The challenge sent to the
  adversary is $(\csig_0,\yeval) \gets \Sign(\PRVUK[\cuid_0],\PUBOK[\oid],
  \CRED[\scid_0],\msg,\feval)$, where $\csig_0 = (\pi_0,\cinsp)$, with
  $\pi_0 = \NIZKProve^{\NIZKRel_{\Sign}}(\NIZKcrs_{\Sign},(\msg,\feval,\yeval,
  \cinsp,\lbrace\PUBIK[\scid_{i,0}]\rbrace_{i\in[n]},\PUBOK[\oid]),
  (\PUBUK[\cuid_0],\PRVUK[\cuid_0],\lbrace (\scid_{i,0},\attrs_{i,0},
  \CRED[\scid_{i,0}]) \rbrace_{i\in[n]},\yeval^1,\yinsp,r,r'))$, and $\cinsp =
  \EEnc(\PUBOK[\oid],\yinsp;r')$, for some random $r,r'$ and $n \ge 0$.
  % 
  Further, we build $G_1^0$ from $G_0^0$ by simulating the proof $\pi_0$. That
  is, in $G_1^0$, $\csig_0 = (\pi_0^s,\ceval,\cinsp)$, where $\pi^s_0 =
  \NIZKSim^{\NIZKRel_{\Sign}}(\NIZKcrs_{\Sign},\NIZKtrap,(\msg,\feval,\yeval,
  \cinsp,\lbrace\PUBIK[\scid_{i,0}]\rbrace_{i\in[n]},\PUBOK[\oid]))$. By
  zero-knowledgeness of $\NIZK^{\Sign}$, $G_1^0$ is indistinguishable from
  $G_0^0$.

  Similarly, we consider $G_0^1$ and $G_1^1$. That is, $G_0^1 = \ExpSigAnono$,
  where the challenge
  sent to the adversary is $(\csig_1,\yeval) \gets \Sign(\PRVUK[\cuid_1],
  \PUBOK[\oid],\CRED[\scid_1],\msg,\feval)$, where $\csig_1 = (\pi_1,
  \cinsp)$, with $\pi_1 = \NIZKProve^{\NIZKRel_{\Sign}}(\NIZKcrs_{\Sign},
  (\msg,\feval,\yeval,
  \cinsp,\lbrace\PUBIK[\scid_{i,1}]\rbrace_{i\in[n]},\PUBOK[\oid]),
  (\PUBUK[\cuid_1],\PRVUK[\cuid_1],\lbrace (\scid_{i,1},\attrs_{i,1},
  \CRED[\scid_{i,1}]) \rbrace_{i\in[n]},\yeval^1,\yinsp,r,r'))$, and
  $\cinsp = \EEnc(\PUBOK[\oid],\yinsp;r')$, for random $r,r'$ and $n \ge 0$. As
  before, build $G_1^1$ from $G_0^1$, simulating $\pi_1$. That is, in $G_1^1$,
  $\csig_1 = (\pi_1^s,\ceval,\cinsp)$, where $\pi^s_1 =
  \NIZKSim^{\NIZKRel_{\Sign}}(\NIZKcrs_{\Sign},\NIZKtrap,(\msg,\feval,\yeval,
  \cinsp,\lbrace\PUBIK[\scid_{i,1}]\rbrace_{i\in[n]},\PUBOK[\oid]))$. Again, by
  zero-knowledge of $\NIZK^{\Sign}$, $G_1^1$ is
  indistinguishable from $G_0^1$.

  Note that, at this point, $G_1^1$ and $G_1^0$ are indistinguishable.
  Concretely, the challenge signatures they produce have the structure $\Sig_b =
  ((\pi_b,\cinsp),\yeval)$ where, to prevent trivial wins, \yeval is enforced
  (in the \SIGN oracle) to be the same for both challenge users, and \cinsp
  either encrypts the same value for both users, or \adv~cannot query the \OPEN
  oracle. \SIGN also enforces that $\PUBIK[\scid_0] = \PUBIK[\scid_1]$. Finally,
  the proofs $\pi_b$ are simulated, and thus reveal no information about the
  signer keys or credentials.

@@

\begin{proof}[\thmref{thm:issue-forge-uas}; Issuance unforgeability of \CUASGen]
  We show that the probability that \fissue outputs a wrong $\yinsp$ value is
  negligible, as well as the probability that the credential identifier does not
  belong to an existing (honest or corrupt) user.

  Observe that the adversary is required to output a credential identifier for
  which associated entries in \trans and \CRED exist; moreover, if such a
  credential was produced by an issuer, we must have access to those entries, as
  issuers are assumed to be honest in this game.
  %
  Since we can access $\trans[\cid] = (\yissue,\utrans)$, and \utrans contains
  a NIZK proof $\NIZKproof_{\Issue}$ proving knowledge by the user of a witness
  in $\NIZKRel_{\Issue}$, which is simulation extractable, \ExtractIssue
  produces a tuple $(\upk,\usk,\lbrace (\cid_i,\attrs_i,\cred_i)
  \rbrace_{i\in[n]}$.
  %
  From the fact that issuers are honest, we know that $\ATTR[\cid]$ must match
  the attribute set \attrs used by user and issuer in the issuance protocol.
  Thus, from  soundness of the NIZK system, the $(\upk,\lbrace (\cid_i,\attrs_i,
  \cred_i) \rbrace_{i\in[n]}$ values within the extracted tuple, along with
  $\ATTR[\cid] = \attrs$ must make \fissue output \yissue, except with
  negligible probability. Moreover, soundness of the NIZK system also implies
  that all credentials are bound to the same \usk, which is associated to the
  extracted \upk.

  Finally, suppose that the extracted \upk does not belong to a \uid in
  $\HU \cup \CU$. Then, the produced credential is a forgery. \todo{How to
    extract the forgery in this case?? I think that expanding the definition
    so that it requires \adv~to produce a valid signature would make things
    much easier...}
  \qed
\end{proof}

@@

\begin{proof}[\thmref{thm:sign-forge-uas}; Signature unforgeability of \CUASGen]
  We show that an adversary \adv~winning \ExpForgeSign can be used to break the
  one-more unforgeability property of \SBCM.

  If the verification at line 4 of \ExpForgeSign holds, then $(\msg,\feval,
  \yeval,\cinsp,\sipk,\opk,\widetilde{\Eek}) \in \NIZKLang^{\Sign}$. Then:
  %

  \paragraph{(a) \Judge accepts $(\yinsp,\iproof)$.} %
  Simulation extractability of $\NIZK^{\Sign}$ thus ensures that $\cinsp =
  \EEnc(\Eek,\yinsp)$, and since $(\yinsp,\iproof)$ is generated honestly at
  line 5, then \yinsp is the correct decryption of \cinsp, and  correctness of
  $\NIZK^{\Open}$ ensures that \Judge outputs $1$ at line 6.

@@

  \begin{itemize}
  \item $\yeval^0 = \yeval$, where $(\yeval^0,\cdot) = \feval(\upk,\lbrace
    (\cid_i,\attrs_i) \rbrace_{i\in[n]},\msg)$, and $\yeval$ is as output by
    \adv~at step 2.    
  \item $\tyeval^1 = \yeval^1$, where $(\cdot,\yeval^1) = \feval(\upk,\lbrace
    (\cid_i,\attrs_i)\rbrace_{i\in[n]},\msg)$, and $\tyeval^1$ is as extracted
    by \ExtractSign at line 7.
  \end{itemize}

  Thus, the probability of \adv~winning at line 9 is negligible, due to
  simulation extractability of $\NIZK^{\Sign}$.

  \paragraph{(c) The output of \finsp matches the output of \Open.} %
  After (a), the \yinsp value output by \Open is the correct decryption of
  \cinsp. After (b), the $(\yeval^0=\yeval,\yeval^1)$ values output by
  \ExtractSign match the evaluation of \feval. Thus, simulation extractability
  of $\NIZK^{\Sign}$ ensures that $\finsp((\yeval^0,\yeval^1),\upk,\lbrace
  (\cid_i,\attrs_i)\rbrace_{i\in[n]},\msg) = \yinsp$ and, also, that \yinsp
  matches the $\yinsp'$ value extracted by \ExtractSign. Thus, the probability
  of \adv~winning at line 10 is negligible due to simulation extractability
  of $\NIZK^{\Sign}$.

  \paragraph{(d) All {\cred}s are bound to the same \usk.} %
  $\NIZKRel^{\Sign}$ includes a condition that $\forall i\in[n],
  \SBCMVerify(\ipk_i,\cred_i,\usk,\attrs_i) = 1$. Thus, simulation
  extractability of $\NIZK^{\Sign}$ and correctness of \SBCM, ensure that all
  credentials involved in the signature contain \usk as their user key (first)
  attribute. Consequently, \IdentifyCred returns $1$ for all the involved
  credentials, and the probability of \adv~winning at line 11 is $0$.
